---
title: "Aides Drug"
author: "Ashley Word Ycaza"
date: "May 31"
output:
  html_document:
    number_sections: true
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
avast <-read.csv(file="avastivironal.csv",header=TRUE)
head(avast)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides (abacavirinol) is under test in clinical trials. Analyze the effectiveness of this drug across the board. Consider population subgroups by gender, drug use, and sexual orientation. 

the first question is this: does abacavirinol have any effect on whether one contracts aids or not?

$H_0:$ abacavirinol has no effect on whether one contracts aids or not  
$H_1:$ abacavirinol has an effect on whether one contracts aids or not  

the remaining questions are: does abacavirinol have a different effect based on gender?

$H_0:$ gender has no effect on whether abacavirinol effectd against contracting aids  
$H_1:$ gender does effect whether abacavirinol has an effect on contracting aids  

does abacavirinol have a different effect based on drug use?  

$H_0:$ drug use has no effect on whether abacavirinol has an effect on contracting aids  
$H_1:$ drug use does have an effect on whether abacavirinol has an effect on contracting aids  

does abacavirinol have a different effect based on sexual orientation?

$H_0:$ sexual orientaion has no effect on whether abacavirinol has an effect on contracting aids  
$H_1:$ sexual orientation does have an effect on whether abacavirinol has an effect on contracting aids  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

for each hypothesis test, the graphical summary will be a bar chart, the numerical summary will be a cross tabulation, and a chi-squared test will be used for the inferential results

#Results General Effectiveness

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

Is the vaccine effective on the general population  


## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~treatment+infected, data=avast, type= "percent")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
genTable = xtabs(~treatment+infected, data=avast)
print(genTable)
```


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(genTable)
```


#Results Gender effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

does gender have an effect on abacavirinol effectiveness  

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~sex+infected, data=avast, type= "percent")
```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
genTable = xtabs(~sex+infected, data=avast)
print(genTable)
```


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(genTable)
```


#Results Drug Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

does drug use have an effect on abacavirinol effectiveness  

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~Ivdrug+infected, data=avast, type= "percent")

```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
genTable = xtabs(~Ivdrug+infected, data=avast)
print(genTable)
```


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(genTable)
```


#Results Sexual Orientation Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

does sexual orientation have an effect on abacavirinol effectiveness  

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~LBGT+infected, data=avast, type= "percent")
```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
genTable = xtabs(~LBGT+infected, data=avast)
print(genTable)
```


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(genTable)
```



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

abacavirinol has no effect on whether one contracts aids or not

gender does effect whether abacavirinol has an effect on contracting aids  

drug use has no effect on whether abacavirinol has an effect on contracting aids  

sexual orientaion has no effect on whether abacavirinol has an effect on contracting aids

